Evaluating the evidence of widespread maintenance of functional composition in vertebrate communities

### 2022-04-28 paper/manuscript.Rmd
final stuff

It is critical to understand how ecological communities are responding to the ongoing biodiversity crisis. Previous broad-scale assessments of community change through time have focused on metrics of species identity, which do not account for differences between species. Functional trait approaches offer a powerful alternative to species-based approaches by directly assessing changes in the species characteristics that define their ecological role. Here, we pair vertebrate community time series with functional trait data to take a first look at temporal trends in functional composition across biomes, realms, and taxa. We found no evidence of systematic functional loss across communities, with maintenance of all measured aspects of functional structure. We also identified key methodological holes that may be obscuring a potential underlying trend.

@@

We thank the scientists who contributed to and maintain the biodiversity databases included in this study, including BioTIME, Amphibio, and Elton Traits.
We thank Dr Shane Blowes and Dr Sarah Supp, whose code we adapted for initial data processing of the BioTIME database.
Support for this work was provided by U.S. Department of Energy through the Computational Sciences Graduate Fellowship (DE-FG02-97ER25308) awarded to K.E.A.N..

>>>>

How are ecological communities responding to anthropogenic change? Functional trait approaches allow us to assess this questions by directly measuring shifts in the characteristics that define a species' ecological role. Here, we pair vertebrate community time series with functional trait data to take a first look at temporal trends in functional composition across biomes, realms, and taxa. We found no evidence of systematic functional loss across communities, with maintenance of all measured aspects of functional structure. We also highlight the need for targeted data collection and methodological expansion to further assess functional trends.

@@

We thank the scientists who contributed to and maintain the biodiversity databases included in this study, including BioTIME, Amphibio, and Elton Traits.
We thank Dr Shane Blowes and Dr Sarah Supp, whose code we adapted for initial data processing of the BioTIME database.
Support for this work was provided by U.S. Department of Energy through the Computational Sciences Graduate Fellowship (DE-FG02-97ER25308) awarded to KEAN. CB acknowledges support from NSF CAREER Award #1942280 & NIFA project CA-B-INS-0162-H.
